---
title: "Estimation of sugar accumulation time"
author: "Flora TAVERNIER"
date: "`r substr(Sys.time(), 1, 10)`"
output:
  rmdformats::downcute
---

```{r rm tout, echo=FALSE, eval=FALSE}
rm(list=ls())
```

```{r setup, echo=TRUE, message=FALSE, results='hide', cache=FALSE}
knitr::opts_chunk$set(
	message = FALSE,
	warning = FALSE,
	cache.lazy = FALSE
)
library(anyLib)
anyLib(c("readxl", "readODS", "tidyverse", "janitor", "apercu", "signal", 
         "kableExtra", "gridExtra", "ggplot2", "ggsignif", "ggpubr", 
         "pls", "psych", "MASS", "prospectr", "rchemo", "ChemoSpec", #"roboost", 
         "heatmaply", "factoextra", "ade4", "FactoMineR", 
         "RColorBrewer", "colorRamps", "viridis"))
```

# Loading data and functions

```{r}
load("Env_tmps_acc_sucres_baies.RData")
```

# Data preparation

## NIRS monitoring

```{r}
nirs.df <- nirs.list.monit$imp
```

```{r}
# Change date format
nirs.df$date_sp <- as.Date(nirs.df$date_sp, format = "%Y-%m-%d")
nirs.df$date_sp <- format(nirs.df$date_sp, format = "%d.%m.%y")

# New set of rownames for NIRS
nirs.df$echantillon <- paste(nirs.df$echantillon, nirs.df$date_sp, sep = ".")
nirs.df$echantillon <- gsub("-", ".", nirs.df$echantillon)
nirs <- as.data.frame(nirs.df$NIRS)
rownames(nirs) <- 1:nrow(nirs)
```

## Veraison prediction

```{r}
veraison_pred_monit <- predict(mod_plsda, nirs)
```

```{r}
veraison_pred_monit.df <- as.data.frame(veraison_pred_monit)

veraison_pred_monit.df$echantillon <- nirs.df$echantillon

colnames(veraison_pred_monit.df) <- 
  c("pred", colnames(veraison_pred_monit.df[, 2:ncol(veraison_pred_monit.df)]))
```

## Select berries after veraison only

```{r}
# baiesV <- subset(veraison_pred_monit.df[, c(1, 5)], veraison_pred_monit.df$pred == "V")

baiesV <- veraison_pred_monit.df[, c(1, 5)]
```

## Merge monitoring informations

```{r}
baiesV <- baiesV
pred.ok.noout.ok <- pred.ok.noout

# Convert the ‘date_sp’ column to Date format
pred.ok.noout.ok$date_sp <- as.Date(pred.ok.noout.ok$date_sp, format = "%Y-%m-%d")

# Change date format to dd.mm.yy
pred.ok.noout.ok$date_sp <- format(pred.ok.noout.ok$date_sp, format = "%d.%m.%y")

# Sample creation with bay ID and date
pred.ok.noout.ok$echantillon <- paste(pred.ok.noout.ok$grappebaie, pred.ok.noout.ok$date_sp, sep = ".")

# Replace ‘-’ with ‘.’ in the ‘sample’ column
pred.ok.noout.ok$echantillon <- gsub("-", ".", pred.ok.noout.ok$echantillon)

# Merge
baiesV.ok <- merge(baiesV, pred.ok.noout.ok, by = "echantillon")
```

```{r}
# Drop duplicates
length(which(duplicated(baiesV.ok$echantillon) == "TRUE"))
dup <- baiesV.ok[which(duplicated(baiesV.ok$echantillon) == "TRUE"),]
```
```{r}
# Change data format
baiesV.ok <- as.data.frame(baiesV.ok)
baiesV.ok[] <- lapply(baiesV.ok, function(x) if(is.list(x)) as.character(x) else x)
```

```{r}
baiesV.ok$UniqueID <- paste(baiesV.ok$echantillon, baiesV.ok$num, sep = "_")

# Delete duplicates by keeping the first occurrence
baiesV.ok <- baiesV.ok[!duplicated(baiesV.ok$UniqueID), ]
```

```{r}
baiesV.ok <- baiesV.ok %>% relocate(UniqueID, .before = echantillon) %>% 
  relocate(num, .after = echantillon) %>% 
  relocate(pred, .after = DateDOY) %>% 
  relocate(sum.Gpred.Fpred, .after = Notes)

# baiesV.ok <- baiesV.ok[, -c(3, 9, 12, 13, 15:17)]

baiesV.ok <- baiesV.ok %>% rename(sugar_cal = sum.Gpred.Fpred)
```

```{r}
baiesV.ok$DateDOY <- as.numeric(baiesV.ok$DateDOY)
```

# Polynomial fit for each berry

## Sorting healthy berries (without notes)

```{r}
print("Account with notes")
nrow(baiesV.ok)

# Filter
df_sain <- baiesV.ok[is.na(baiesV.ok$Notes), ]

print("Account whitout notes")
nrow(df_sain)
```

## Sorting berries > 1.5 M G+F

```{r}
print("Account all G+F")
nrow(df_sain)

# Filter
df_sain$sugar_cal[which(df_sain$sugar_cal < 0)] <- 0
df_sain <- df_sain[df_sain$sugar_cal <= 1500, ]


print("Account of berries between 0 and 1500 mM G+F")
nrow(df_sain)
```

## Select 2022 genotypes

```{r}
cep2022 <- c("Ugni blanc", "Mourvèdre", "Morrastel", "Grenache", "Riesling", "Carmenère")
subset_df_sain <- df_sain[grep(paste(cep2022, collapse = "|"), df_sain$geno), ]

# Replace special characters
subset_df_sain$geno <- gsub("è", "e", subset_df_sain$geno)
```

## Polynomial: order 2

```{r}
fitpoly2 <- NULL
poly_all2 <- NULL
for (genot in unique(subset_df_sain$geno)) {
  # Subset genotype
  sub.geno <- subset(subset_df_sain, subset_df_sain$geno == genot)
  # Select berries with at least 9 dates
  b9dat <- names(which(table(sub.geno$genoID) >= 9))
  # The date 200 has been removed because the spectra had a problem on that date
  dat <- subset(sub.geno, genoID %in% b9dat & doy != 200)

  for (b in unique(dat$grappebaie)) {
    # Subset berry
    sub <- subset(dat, dat$grappebaie == b)
    
    # Fitting a polynomial model with lm()
    sub$DateDOY <- as.numeric(sub$DateDOY)
    fit2 <- lm(sugar_cal ~ stats::poly(DateDOY, 2, raw = T), data = sub)
    sub$polynomial_sugar_cal <- predict(fit2, newdata = sub)
    
    # Plot
    p <- 
      ggplot(sub, aes(x = DateDOY, y = sugar_cal)) +
      geom_line() +
      geom_point() +
      geom_line(aes(y = polynomial_sugar_cal), # Predicted polynomial curve
                color = "blue", 
                linetype = "dashed") + 
      geom_hline(yintercept = 1000, color = "darkorchid") +
      geom_hline(yintercept = 200, color = "forestgreen") +
      labs(x = "Day of Year (DOY)", y = "Sugar_cal (G pred + F pred)",
           title = paste("Polynomial 2", genot, b)) +
      theme_minimal()
    
    print(p)
    
    # Save polynomial
    poly_all2[[paste(genot, b, sep = "_")]] <- sub
    fitpoly2[[paste(genot, b, sep = "_")]] <- fit2
  }
}
```

# Logistic regression (sigmoid fit) for each berry

```{r fig.height=4, fig.width=6, message=TRUE, warning=TRUE}
fitlogi <- NULL
logi_all <- NULL
coef_logi <- NULL
summR2_logi <- NULL

for (genot in unique(subset_df_sain$geno)) {
  sub.geno <- subset(subset_df_sain, subset_df_sain$geno == genot)
  b9dat <- names(which(table(sub.geno$genoID) >= 9))
  dat <- subset(sub.geno, genoID %in% b9dat & doy != 200)
  
  for (b in unique(dat$grappebaie)) {
    sub <- subset(dat, dat$grappebaie == b)
    sub$DateDOY <- as.numeric(sub$DateDOY)
    
    # Normalisation of data by their max
    # sub$sugar_cal <- sub$sugar_cal / max(sub$sugar_cal, na.rm = TRUE)
    
    # Fitting a logistic regression with nls() by providing explicit initial values:
    tryCatch({
      fit <- nls(sugar_cal ~ SSlogis(DateDOY, Asym, xmid, scal),
                 data = sub,
                 # make the convergence algorithm more tolerant:
                 control = nls.control(minFactor = 1e-8))
    # fit <- nls(sugar_cal ~ SSlogis(DateDOY, Asym, xmid, scal), data = sub)
    sub$reg_logi_sugar_cal <- predict(fit, newdata = sub)
    
    # R² for logistic regression
    # Sum of total squares:
    ss_total <- sum((sub$sugar_cal - mean(sub$sugar_cal)) ^ 2)
    # Sum of residual squares:
    ss_res <- sum(residuals(fit) ^ 2)
    # Calculation of pseudo-R²:
    r2 <- 1 - (ss_res / ss_total)
    
    # Plot
    p <- 
      ggplot(sub, aes(x = DateDOY, y = sugar_cal)) +
      geom_line() +
      geom_point() +
      geom_line(aes(y = reg_logi_sugar_cal), # Predicted sigmoid curve
                color = "blue", 
                linetype = "dashed") + 
      geom_hline(yintercept = 1000, color = "darkorchid") +
      geom_hline(yintercept = 200, color = "forestgreen") +
      labs(x = "Day of Year (DOY)", y = "Sugar_cal (G pred + F pred)",
           title = paste("Logistic regression", genot, b)) +
      theme_minimal()
    
    print(p)
    
    # Save sigmoid fit
    logi_all[[paste(genot, b, sep = "_")]] <- sub
    fitlogi[[paste(genot, b, sep = "_")]] <- fit
    coef_logi[[paste(genot, b, sep = "_")]] <- coef(fit)
    summR2_logi[[paste(genot, b, sep = "_")]] <- r2
    },
    error = function(e) { # Error message
      message(paste("Adjustment error for", genot, b, ":", e$message))
    })
  }
}
```

## Output polynomial

### Dataframe of predicted sugar_cal values

```{r}
poly_all_df <- do.call(rbind, poly_all2)
```

```{r}
poly_all_df.ok <- poly_all_df
```

```{r}
poly_all_df.ok <- poly_all_df.ok[, c("geno", "grappebaie", "doy", "polynomial_sugar_cal", "glucose", "fructose", "malate", "sugar_cal", "poids_baie", "Notes")]
```

### Polynomial fit output

```{r}
fitpoly2.ok <- NULL
for (i in names(fitpoly2)) {
    fitpoly2.ok[[i]] <- fitpoly2[[i]][[1]]
}
```

```{r}
fitpoly2.coef <- as.data.frame(do.call(rbind, fitpoly2.ok))

fitpoly2.coef$echantillon <- rownames(fitpoly2.coef)

colnames(fitpoly2.coef) <- c("c", "b", "a", "echantillon")
```

```{r}
cep2022 <- c("Ugni blanc", "Mourvèdre", "Morrastel", "Grenache", "Riesling", "Carmenère")

subset_fitpoly2 <- fitpoly2.coef[grep(paste(cep2022, collapse = "|"), fitpoly2.coef$echantillon), ]
```

### Summary of polynomial fit

```{r}
summi <- NULL
for (i in names(fitpoly2)) {
  # print(i)
  summi[[i]] <- summary(fitpoly2[[i]])
  # print(summi[[i]])
}
```

```{r}
r_squared_values <- sapply(summi, function(x) x$r.squared)
names_summi <- names(summi)
summi.df <- data.frame(echantillon = names_summi, R_squared = r_squared_values)

subset_summi.df <- summi.df[grep(paste(cep2022, collapse = "|"), summi.df$echantillon), ]

coeff.rsq.poly2 <- merge(subset_summi.df, subset_fitpoly2, by = "echantillon")
coeff.rsq.poly2$echantillon <- gsub("è", "e", coeff.rsq.poly2$echantillon)
```

```{r}
coeff.rsq.poly2$baieid <- coeff.rsq.poly2$echantillon
coeff.rsq.poly2$baieid <- gsub("_", "", coeff.rsq.poly2$baieid)
coeff.rsq.poly2$baieid <- gsub("-", "baie", coeff.rsq.poly2$baieid)
coeff.rsq.poly2$baieid <- gsub(" ", "", coeff.rsq.poly2$baieid)

poly_all_df.ok$geno <- gsub("è", "e", poly_all_df.ok$geno)
```

### Save outputs

```{r eval=FALSE}
# Polynomial data:
setwd("D:/Mes Donnees/Doctorat/Axe 1 - NIRS/Model_final_propre_OK")
write.csv2(poly_all_df.ok, file = "Data_AfterVeraison_for_calcul_time.csv")

write.csv2(poly_all_df.ok, file = "Data_for_calcul_time_poly2.csv")
```

```{r eval=FALSE}
# Dataframe of filtered data (whitout notes ans only 2022):
setwd("D:/Mes Donnees/Doctorat/Axe 1 - NIRS/Model_final_propre_OK")
write.csv2(subset_df_sain, file = "Data_for_calcul_time.csv")
```

```{r eval=FALSE}
# Coefficients of the polynomial and R²:
setwd("D:/Mes Donnees/Doctorat/Axe 1 - NIRS/Model_final_propre_OK")
write.csv2(coeff.rsq.poly2, file = "Data_for_calcul_time_rsquare_poly2.csv")
```

## Output logistic regression (sigmoid)

### Dataframe of predicted sugar_cal values

```{r}
logi_all_df <- as.data.frame(do.call(rbind, logi_all))
```

```{r}
logi_all_df.ok <- logi_all_df
```

```{r}
logi_all_df.ok <- logi_all_df.ok[, c("geno", "grappebaie", "doy", "reg_logi_sugar_cal", "glucose", "fructose", "malate", "sugar_cal", "poids_baie", "Notes")]
```

### Sigmoid fit output

```{r}
coef_logi.ok <- as.data.frame(do.call(rbind, coef_logi))
coef_logi.ok$echantillon <- rownames(coef_logi.ok)

summR2_logi.df <- as.data.frame(do.call(rbind, summR2_logi))
summR2_logi.df$echantillon <- rownames(summR2_logi.df)
```

```{r}
coef_R2_logi <- merge(coef_logi.ok, summR2_logi.df, by = "echantillon")
```

```{r}
logi_all_df.ok$geno <- gsub("è", "e", logi_all_df.ok$geno)
```

### Save outputs

```{r eval=FALSE}
# Sigmoid data:
setwd("D:/Mes Donnees/Doctorat/Axe 1 - NIRS/Model_final_propre_OK")
write.csv2(logi_all_df.ok, file = "Data_for_calcul_time_logi.csv")
```

```{r eval=FALSE}
# Coefficients of the sigmoid + R²:
setwd("D:/Mes Donnees/Doctorat/Axe 1 - NIRS/Model_final_propre_OK")
write.csv2(coef_R2_logi, file = "coef_logi_rsq.csv")
```

# Examination of the R²

```{r}
# Selection of the same berries for the two fits
rsq_poly_sigm <- merge(coeff.rsq.poly2, coef_R2_logi, by = "echantillon")

colnames(rsq_poly_sigm) <- c("echantillon", 
                             "R²_poly2", 
                             "c", 
                             "b", 
                             "a", 
                             "baieid", 
                             "Asym", 
                             "xmid", 
                             "scal",
                             "R²_sigmoid")
```

## Histogram of the R² values

```{r}
hist(rsq_poly_sigm$`R²_poly2`, 
     breaks = 30, 
     xlim = c(0.5, 1),
     ylim = c(0, 15),
     col = rgb(1, 0, 0, 0.5), 
     xlab = "R²", ylab = "Frequency", 
     main = "Distribution of R² for Polynomial 2 and Sigmoid")

hist(rsq_poly_sigm$`R²_sigmoid`, 
     breaks = 30, 
     xlim = c(0.5, 1),
     ylim = c(0, 15),
     col = rgb(0, 0, 1, 0.5),
     add = T)

legend("topleft", legend = c("Poly2","Sigmoid"), col = c(rgb(1, 0, 0, 0.5), 
     rgb(0, 0, 1, 0.5)), pt.cex = 2, pch = 15, bty = "n")
```

```{r}
# Defining the layout of graphics
par(mfrow = c(1, 2), mar = c(4, 4, 2, 0), oma = c(0, 0, 2, 0)) # oma= adds space for a common title

# Polynomial hist
hist(rsq_poly_sigm$`R²_poly2`, 
     breaks = 20, 
     xlim = c(0.5, 1),
     ylim = c(0, 20),
     col = rgb(1, 0, 0, 0.5), 
     xlab = "R²", ylab = "Frequency", 
     main = "")

legend("topleft", legend = c("Poly2","Sigmoid"), col = c(rgb(1, 0, 0, 0.5), 
     rgb(0, 0, 1, 0.5)), pt.cex = 2, pch = 15, bty = "n")

# Sigmoid hist
hist(rsq_poly_sigm$`R²_sigmoid`, 
     breaks = 20, 
     xlim = c(0.5, 1),
     ylim = c(0, 20),
     col = rgb(0, 0, 1, 0.5), 
     xlab = "R²", ylab = "", 
     main = "")

# Common title
mtext("Distribution of R² for Polynomial 2 and Sigmoid", 
      outer = TRUE, cex = 1.5)
```

## Wilcoxon test between fitting methods

```{r}
wilcox_test_result <- wilcox.test(rsq_poly_sigm$`R²_poly2`,
                                  rsq_poly_sigm$`R²_sigmoid`,
                                  paired = T)
print(wilcox_test_result)
```

```{r}
data <- data.frame(
  value = c(rsq_poly_sigm$`R²_poly2`, rsq_poly_sigm$`R²_sigmoid`),
  group = factor(rep(c("R²_poly2", "R²_sigmoid"), each = length(rsq_poly_sigm$`R²_poly2`)))
)
```

<!-- ```{r eval=FALSE} -->
<!-- ggplot(data, aes(x = group, y = value, fill = group)) + -->
<!--   # Display the complete violin shape -->
<!--   geom_violin(trim = F, color = NA, alpha = 0.5) + -->
<!--   geom_boxplot(width = 0.1) + -->
<!--   xlab("") + ylab("R² Values") +  -->
<!--   ggtitle("Violin Plot of R² for Polynomial 2 and Sigmoid") + -->

<!--   # Add individual data points and limit overlap -->
<!--   # geom_jitter(aes(color = group), width = 0.2, size = 1.5) + -->
<!--   # Add mean -->
<!--   stat_summary(fun = "mean",  -->
<!--                geom = "point",  -->
<!--                shape = 17,  -->
<!--                size = 3,  -->
<!--                color = "black") + -->
<!--   theme_minimal() + -->
<!--   theme(legend.position = "none", -->
<!--         plot.title = element_text(size = 18, face = "bold"),  -->
<!--         axis.title.y = element_text(size = 16),  -->
<!--         axis.title.x = element_text(size = 16),  -->
<!--         axis.text.x = element_text(hjust = 1, size = 14),  -->
<!--         axis.text.y = element_text(hjust = 1, size = 14)) + -->

<!--   # Add statistical test comparisons with geom_signif() -->
<!--   geom_signif(comparisons = list(c("R²_poly2", "R²_sigmoid")),  -->
<!--               map_signif_level = T,  -->
<!--               test = "wilcox.test",  -->
<!--               step_increase = 0.1) -->
<!-- ``` -->

```{r}
ggpaired(data, x = "group", y = "value",
         color = "group", line.color = "gray", line.size = 0.4,
         palette = "jco")+
  stat_compare_means(paired = TRUE)
```

## Test by variety

```{r}
rsq_poly_sigm <- coef_R2_logi %>%
  separate(echantillon, into = c("geno", "grappe", "baies"), sep = "_|-")
```

```{r}
print("Number of berries:")
nrow(rsq_poly_sigm)

print("Number of berries per genotype:")
counts <- as.data.frame(table(rsq_poly_sigm$geno))
counts
```

```{r}
ggplot(rsq_poly_sigm, aes(x = geno, y = V1, fill = geno)) +
  geom_violin(trim = F, color = NA, alpha = 0.5) +
  geom_boxplot(width = 0.1) +
  xlab("") + ylab("R² Values") + 
  ggtitle("R² for each variety") +
  stat_summary(fun = "mean", 
               geom = "point", 
               shape = 17, 
               size = 3, 
               color = "black") +
  theme_minimal() +
  theme(legend.position = "none",
        plot.title = element_text(size = 18, face = "bold"), 
        axis.title.y = element_text(size = 16), 
        axis.title.x = element_text(size = 16), 
        axis.text.x = element_text(size = 14, hjust = 1, angle = 45), 
        axis.text.y = element_text(size = 14)) +

  scale_x_discrete(labels = function(x) {
    paste0(x, "\n(n = ", counts[match(x, counts$Var1), "Freq"], ")")})
```

## Filter R² > 0.8 for the sigmoid

```{r}
rsq_poly_sigm.08 <- subset(rsq_poly_sigm, rsq_poly_sigm$V1 >= 0.8)
```

```{r}
print("Number of berries:")
nrow(rsq_poly_sigm.08)

print("Number of berries per genotype:")
counts08 <- as.data.frame(table(rsq_poly_sigm.08$geno))
counts08
```

```{r}
ggplot(rsq_poly_sigm.08, aes(x = geno, y = V1, fill = geno)) +
  geom_violin(trim = F, color = NA, alpha = 0.5) +
  geom_boxplot(width = 0.1) +
  xlab("") + ylab("R² Values") + 
  ggtitle("R² > ou = à 0.8 for each variety") +
  stat_summary(fun = "mean", 
               geom = "point", 
               shape = 17, 
               size = 3, 
               color = "black") +
  theme_minimal() +
  theme(legend.position = "none",
        plot.title = element_text(size = 18, face = "bold"), 
        axis.title.y = element_text(size = 16), 
        axis.title.x = element_text(size = 16), 
        axis.text.x = element_text(size = 14, hjust = 1, angle = 45), 
        axis.text.y = element_text(size = 14)) +

  scale_x_discrete(labels = function(x) {
    paste0(x, "\n(n = ", counts08[match(x, counts$Var1), "Freq"], ")")})
```
