---
title: "Calibrations/Validation NIRS berry - 2021 + 2022"
author: "F Tavernier, V Segura"
date: "`r substr(Sys.time(), 1, 10)`"
output:
  rmdformats::downcute
---

```{r rm tout, echo=FALSE}
rm(list=ls())
```

```{r setup, echo=TRUE, message=FALSE, results='hide', cache=FALSE}
knitr::opts_chunk$set(echo = TRUE, cache.lazy = FALSE, dev.args = list(bg = "transparent"))
library(anyLib)
anyLib(c("apercu", "RColorBrewer", "signal", "prospectr", "ade4", "factoextra", "readxl", "viridis", "psych", "stringr", "pls", "dplyr", "kableExtra", "MASS", "tidyr", "readODS", "janitor", "colorRamps"))

source("G:/Mon Drive/Thèse MétabEAU/Axe 1 - NIRS/Pred_modele_Elias/predictions/calibSPIR.R")
```

# Preparation of the training and validation sets

## Loading data

### Spectra and HPLC data

```{r}
# Load nirs.df.all
load("NIRSBaies_2021et2022_HPLC.RData")
```

### General information about the dataset

Number of spectra: `r nrow(nirs.df.all)`

Number of cepages: `r length(unique(nirs.df.all$cépage))`

```{r echo=FALSE, message=FALSE, warning=FALSE}
print("Number of varieties")
unique(nirs.df.all$cépage)
```

Number of dates: 

```{r echo=FALSE, message=FALSE, warning=FALSE}
# All dates
levels(unique(as.factor(nirs.df.all$date_sp)))
levels(unique(as.factor(nirs.df.all$DateDOY)))
length(unique(nirs.df.all$date_sp))
```

```{r echo=FALSE, message=FALSE, warning=FALSE}
# Dates per year
levels(unique(as.factor(nirs.df.all$date_sp[which(nirs.df.all$millesime == "2021")])))
length(unique(nirs.df.all$date_sp[which(nirs.df.all$millesime == "2021")]))

levels(unique(as.factor(nirs.df.all$date_sp[which(nirs.df.all$millesime == "2022")])))
length(unique(nirs.df.all$date_sp[which(nirs.df.all$millesime == "2022")]))
```

## Plot raw spectra

```{r fig.height=4, fig.width=6.5}
# 2021
matplot(
  x = as.numeric(colnames(nirs.df.all$NIRS)),
  y = t(subset(nirs.df.all, millesime == "2021")[,"NIRS"]),
  type = "l",
  lty = 1,
  xlab = "Wavelength (nm)",
  ylab = "Absorbance",
  main = "Raw spectra - 2021")

# 2022
matplot(
  x = as.numeric(colnames(nirs.df.all$NIRS)),
  y = t(subset(nirs.df.all, millesime == "2022")[,"NIRS"]),
  type = "l",
  lty = 1,
  xlab = "Wavelength (nm)",
  ylab = "Absorbance",
  main = "Raw spectra - 2022")

# All
matplot(
  x = as.numeric(colnames(nirs.df.all$NIRS)),
  y = t(nirs.df.all$NIRS),
  type = "l",
  lty = 1,
  xlab = "Wavelength (nm)",
  ylab = "Absorbance",
  main = "Raw spectra - 2021 and 2022")

# All mean dates
nirs.mean.date <- apply(nirs.df.all$NIRS, 2, function(x) {tapply(X = x, INDEX = nirs.df.all$date_sp, FUN = function(y) {mean(y, na.rm = TRUE)})})

couleur = viridis::viridis(n = nrow(nirs.mean.date))

par(mar = c(4.5, 4.5, 3, 8))
matplot(x = as.numeric(colnames(nirs.mean.date)), y = t(nirs.mean.date),
  type = "l", lty = 1, col = couleur, xlab = "Wavelength (nm)", ylab = "Absorbance", main = "Mean spectra - by date")
legend(x = 1730, y = 0.55, legend = rownames(nirs.mean.date), lty = 1, col = couleur, bty = "n", xpd = TRUE, cex = 0.8)
```

## Preprocessing

```{r}
# Imputation
nirs.imp <- nirs.df.all
nirs.imp$NIRS <- t(apply(nirs.df.all$NIRS, 1, function(x) {
  x[is.na(x)] <- max(x, na.rm = TRUE)
  return(x)
}))
matplot(x = as.numeric(colnames(nirs.imp$NIRS)), y = t(nirs.imp$NIRS), type = "l", lty = 1,
        xlab = "Wavelength (nm)", ylab = "Absorbance", main = "NAs imp.")

# SNV
nirs.snv <- nirs.imp
nirs.snv$NIRS <- t(scale(t(nirs.imp$NIRS)))
matplot(x = as.numeric(colnames(nirs.snv$NIRS)), y = t(nirs.snv$NIRS), type = "l", lty = 1,
        xlab = "Wavelength (nm)", ylab = "Absorbance", main = "SNV")

# SNV no scale
nirs.snv.cent <- nirs.imp
nirs.snv.cent$NIRS <- t(scale(t(nirs.imp$NIRS), center = T, scale = F))
matplot(x = as.numeric(colnames(nirs.snv.cent$NIRS)), y = t(nirs.snv.cent$NIRS), type = "l", lty = 1,
        xlab = "Wavelength (nm)", ylab = "Absorbance", main = "SNV")

# Detrend
nirs.dt <- nirs.imp
nirs.dt$NIRS <- detrend(X = nirs.imp$NIRS, wav = as.numeric(colnames(nirs.imp$NIRS)))
matplot(x = as.numeric(colnames(nirs.dt$NIRS)), y = t(nirs.dt$NIRS), type = "l", lty = 1,
        xlab = "Wavelength (nm)", ylab = "Absorbance", main = "Detrend")

# Derivative
tsf <- (max(as.numeric(colnames(nirs.snv$NIRS))) - min(as.numeric(colnames(nirs.snv$NIRS)))) /(length(as.numeric(colnames(nirs.snv$NIRS))) - 1)
# Der1
nirs.der1 <- nirs.snv
nirs.der1$NIRS <- t(apply(nirs.snv$NIRS, 1, function(x) {sgolayfilt(x, p = 2, n = 11, m = 1, ts = tsf)}))
colnames(nirs.der1$NIRS) <- colnames(nirs.snv$NIRS)
matplot(x = as.numeric(colnames(nirs.der1$NIRS)), y = t(nirs.der1$NIRS), type = "l", lty = 1,xlab = "Wavelength (nm)", ylab = "Absorbance", main = "Der1")
# Der2
nirs.der2 <- nirs.snv
nirs.der2$NIRS <- t(apply(nirs.snv$NIRS, 1, function(x) {sgolayfilt(x, p = 3, n = 21, m = 2, ts = tsf)}))
colnames(nirs.der2$NIRS) <- colnames(nirs.snv$NIRS)
matplot(x = as.numeric(colnames(nirs.der2$NIRS)), y = t(nirs.der2$NIRS), type = "l", lty = 1,xlab = "Wavelength (nm)", ylab = "Absorbance", main = "Der2")

# nirs.list.all
nirs.list <- list("imp" = nirs.imp, "snv" = nirs.snv, "snv.cent" = nirs.snv.cent, "dt" = nirs.dt, "der1" = nirs.der1, "der2" = nirs.der2)

# Cleaning
rm(nirs.imp, nirs.dt, nirs.snv, nirs.der1, nirs.der2, nirs.snv.cent, nirs.mean.date, tsf)
```

## Data distribution and correlations

```{r fig.height=10, fig.width=10}
# Traits used for calibration
trait.ids <- c("glucose", "fructose", "somme_GF", "rapport_GF", "malate", "tartrate", "shikimate", "rapport_MT", "poids_baie")

# Colour vector
nirs.df.all$cépage <- as.factor(nirs.df.all$cépage)
nirs.df.all$DateDOY <- as.factor(nirs.df.all$DateDOY)
col.cep <- viridis::magma(n = length(nirs.df.all$cépage))
col.doy <- viridis::viridis(n = length(levels(nirs.df.all$DateDOY)))

# Correlation matrix
pairs.panels(x = nirs.df.all[, trait.ids], bg = col.doy[nirs.df.all$DateDOY], pch = 21,
             hist.col = "grey", main = "Temps en jours de l'année depuis le 01/01 (DOY)")
# pairs.panels(x = nirs.df.all[, trait.ids], bg = col.cep, pch = 21,
#              hist.col = "grey", main = "Cépage")
```

## Calib with outlier detection and elimination

```{r cache = TRUE, fig.height=6, fig.width=12}
output.cal.noout <- NULL
output.pred.noout <- NULL
outliers <- NULL
for (i in trait.ids){
  #data prep
  datCal <- lapply(nirs.list, function(x){
    droplevels(na.omit(x[, c("echantillon", "DateDOY", "cépage", i, "NIRS")]))
  })
  #model
  model <- calibrationSPIR(calibSetList = datCal, chem = i, maxcomps = 20,
                           LOO = FALSE, folds = 5, iterMCCV = 50, wold = 1, thresh = 0.001)
  print(model$outcalib)
  #best pretr
  bestpretr <- rownames(model$outcalib)[which.max(model$outcalib$R2.MCCV.mean)]
  print(bestpretr)
  #best pretr mod
  datCal.noout <- subset(datCal[[bestpretr]], ! echantillon %in% model$outliers[[bestpretr]])
  bestpretrmod <- MCCV(dat = datCal.noout, trait = i,
                       maxcomp = 20, fold = 5,
                       iter = 50, crit = 1)
  bestpretrmod$output
  #plots
  datCal.noout$cépage <- as.factor(datCal.noout$cépage)
  par(mfrow = c(1, 2))
  #plot by DOY
  plotMCCV(bestpretrmod, pch = 19,
           col = col.doy[as.factor(datCal.noout$DateDOY)],
           main = i, axe = F)
  axis(side = 1)
  axis(side = 2)
  legend("bottomright", legend = paste0("pretr = ", bestpretr), bty = "n")
  legend("right", legend = levels(as.factor(datCal.noout$DateDOY)), pch = 19, col = col.doy, bty = "n")
  #plot by cépage
  plotMCCV(bestpretrmod, pch = 19,
           col = col.cep,
           main = i, axe = F)
  axis(side = 1)
  axis(side = 2)
  legend("right", legend = unique(datCal.noout$cépage), pch = 19, col = col.cep, bty = "n")
  #output
  out <- data.frame(i, bestpretr, length(bestpretrmod$model$model$trait), length(model$outliers[[bestpretr]]),
                    t(bestpretrmod$output))
  colnames(out) <- c("chem", "pretr", "n", "nb.out", "nb.comp", "R2.train", "R2.MCCV.mean", "R2.MCCV.sd", "RMSE.MCCV.mean", "RMSE.MCCV.sd", "RPD.MCCV.mean", "RPD.MCCV.sd")
  output.cal.noout <- rbind(output.cal.noout, out)
  #outliers
  outliers[[i]] <- model$outliers
  #cleaning
  rm(datCal, model, bestpretr, bestpretrmod, datCal.noout, out)
}
```

Best models:

```{r}
kable(output.cal.noout[, c("chem", "pretr", "n", "nb.out", "nb.comp",
                           "R2.MCCV.mean", "R2.MCCV.sd", "RMSE.MCCV.mean", "RMSE.MCCV.sd")],
      "html", row.names = FALSE, digits = 2) %>%
  kable_styling(bootstrap_options = "striped", full_width = F, font_size = 12)
```

Save outliers:

```{r eval=FALSE}
save(outliers, file = "outliers.RData")
save(nirs.list, file = "NIRSpectra_2021-2022_pretrt.RData")
```

## Separate training and validation sets

### Principle

The data set must be divided into 2 parts: 
+ 4/5ths of the data set + outliers = training set
+ 1/5th of the data set without outliers = validation set

The validation set will be selected from the data set without outliers using the algorithm **Kennard-Stone** of the *prospectr* package (Kennard and Stone, 1969 <https://rdrr.io/cran/prospectr/man/kenStone.html>).

### Retrieving the list of outliers and information

```{r}
# How many outliers in common between all the pretreatments: 1 to 6 = number of pretreatments in common
out.table <- sapply(outliers, function(x) {table(unlist(x))})

# Outliers present in all pre-treatments
out.list <- sapply(out.table, function(x) {names(which(x >= 1))})

# Count by trait
sapply(out.list, length)

# Double table: distribution on the traits
table(table(unlist(out.list)))

# Outliers total count
out.vect <- unique(unlist(out.list))
length(out.vect)

# Determining the count for Kennard-Stone
# Nb whitout outliers:
824 - length(out.vect)
# Nb to select with Kennard-Stone in the set whitout outliers:
824 - length(out.vect) - round(824/5)
# Remaining unselected to be put in the validation set:
round(824/5)
```

### Creation of a list without outliers

```{r}
nirs.list.noout <- lapply(nirs.list, function (x) {
  x[!x$echantillon %in% out.vect, ]
  })
```

<!-- row.names(nirs.list$imp)[which(nirs.list$imp$echantillon %in% out.vect)] -->
<!-- nirs.list.noout <- nirs.list$imp$echantillon[which(nirs.list$imp$echantillon %in% out.vect)] -->

### Kennard-Stone selection

```{r}
# Selection
nirs.list.valid <- lapply(nirs.list.noout, function (x) {
  kenStone(x$NIRS, k = 481, metric = "euclid")
})

# Retrieving data for selection
for (i in names(nirs.list)) {
  nirs.list.test <- lapply(nirs.list.noout, function (x) {
    x[nirs.list.valid$i$model, ]
  })
}

# Retrieving data not selected by kenStone (nirs.list.valid.ok)
for (i in names(nirs.list)) {
  nirs.list.valid.ok <- lapply(nirs.list.noout, function (x) {
  x[!x$echantillon %in% nirs.list.test$i$echantillon, ]
  })
}

# nirs.list.train
for (i in names(nirs.list)) {
nirs.list.train <- lapply(nirs.list, function (x) {
  x[!x$echantillon %in% nirs.list.valid.ok$i$echantillon, ]
  })
}
length(nirs.list.train$imp$echantillon)

# Cleaning
rm(nirs.list.test, nirs.list.valid, out.list, out.table)
```

### Save

```{r eval=FALSE}
setwd("D:/Mes Donnees/Doctorat/Pred_modele_Elias/predictions")
# Rdata
save(list = c("nirs.list.valid.ok", "nirs.list.train"), file = "Valid-Train_Dataset_CalibNIRS_2021-2022.Rdata")

# csv
write.csv2(nirs.list.valid.ok$imp, file = "Validation_dataset_2021-2022_CalibNIRS.csv")
write.csv2(nirs.list.train$imp, file = "Training_dataset_2021-2022_CalibNIRS.csv")
```

# Prediction models

## Loading data

### Spectra and HPLC data

```{r}
setwd("D:/Mes Donnees/Doctorat/Axe 1 - NIRS/Pred_modele_Elias/predictions")
load("NIRSBaies_2021et2022_HPLC_OK.RData") #nirs.df.all : df 2021 + 2022 + col calib/valid
load("NIRSpectra_2021-2022_pretrt.RData") #nirs.list : df all pretreatements
load("Valid-Train_Dataset_CalibNIRS_2021-2022.Rdata") #nirs.list.valid.ok et nirs.list.train
load("outliers.RData")
```

# Calibration with separate training and validation set

## Validation and training set info

Validation set:

```{r cache = TRUE, fig.height=6, fig.width=12}
# Nb individals
nrow(nirs.list.valid.ok$imp)
# Nb cepages
length(unique(nirs.list.valid.ok$imp$cépage))
unique(nirs.list.valid.ok$imp$cépage)
# Dates
unique(nirs.list.valid.ok$imp$date_sp)
levels(unique(as.factor(nirs.list.valid.ok$imp$DateDOY)))
length(unique(nirs.list.valid.ok$imp$date_sp))
# Dates per year
unique(nirs.list.valid.ok$imp$date_sp[which(nirs.list.valid.ok$imp$millesime == "2021")])
length(unique(nirs.list.valid.ok$imp$date_sp[which(nirs.list.valid.ok$imp$millesime == "2021")]))
unique(nirs.list.valid.ok$imp$date_sp[which(nirs.list.valid.ok$imp$millesime == "2022")])
length(unique(nirs.list.valid.ok$imp$date_sp[which(nirs.list.valid.ok$imp$millesime == "2022")]))
```

Training set:

```{r cache = TRUE, fig.height=6, fig.width=12}
# Nb individals
nrow(nirs.list.train$imp)
# Nb cepages
length(unique(nirs.list.train$imp$cépage))
unique(nirs.list.train$imp$cépage)
# Dates
unique(nirs.list.train$imp$date_sp)
levels(unique(as.factor(nirs.list.train$imp$DateDOY)))
length(unique(nirs.list.train$imp$date_sp))
# Dates per year
unique(nirs.list.train$imp$date_sp[which(nirs.list.train$imp$millesime == "2021")])
length(unique(nirs.list.train$imp$date_sp[which(nirs.list.train$imp$millesime == "2021")]))
unique(nirs.list.train$imp$date_sp[which(nirs.list.train$imp$millesime == "2022")])
length(unique(nirs.list.train$imp$date_sp[which(nirs.list.train$imp$millesime == "2022")]))
```

## Setting calib

```{r}
# Traits used for calibration
trait.ids <- c("glucose", "fructose", "somme_GF", "rapport_GF", "malate", "tartrate", "shikimate", "rapport_MT", "poids_baie")

nirs.df.all$cépage <- as.factor(nirs.df.all$cépage)
nirs.df.all$DateDOY <- as.factor(nirs.df.all$DateDOY)
col.cep <- viridis::magma(n = length(nirs.df.all$cépage))
col.doy <- hcl.colors(n = length(levels(nirs.df.all$DateDOY)), palette = "sunset")
```

## Basic calibration with independent validation set

```{r cache = TRUE, fig.height=6, fig.width=12}
output.cal.KS <- NULL
mod.bestpretr <- NULL
for (i in trait.ids){
  #data prep
  datCal <- lapply(nirs.list.train, function(x){
    droplevels(na.omit(x[, c("echantillon", "DateDOY", "cépage", i, "NIRS")]))
  })
  #model
  model <- calibrationSPIR(calibSetList = datCal, chem = i, maxcomps = 20,
                           LOO = FALSE, folds = 5, iterMCCV = 50, wold = 1)
  print(i)
  print(model$outcalib)
  #best pretr
  bestpretr <- rownames(model$outcalib)[which.max(model$outcalib$R2.MCCV.mean)]
  print(bestpretr)
  #best pretr mod
  datCal <- subset(datCal[[bestpretr]], ! echantillon %in% model$outliers[[bestpretr]])
  bestpretrmod <- MCCV(dat = datCal, trait = i,
                       maxcomp = 20, fold = 5,
                       iter = 50, crit = 1)
  bestpretrmod$output
  datVal <- lapply(nirs.list.valid.ok, function(x){
    droplevels(na.omit(x[, c("echantillon", "DateDOY", "cépage", i, "NIRS")]))
    })
  #plots
  par(mfrow = c(1, 2))
  #plot by DOY
  plotMCCV(bestpretrmod, pch = 19,
           col = col.doy[as.factor(datCal$DateDOY)],
           main = paste(i, "(cv)", sep = " "), axe = F, col.main = "black", 
           col.lab = "black")
  axis(side = 1, col.axis = "black", col = "black")
  axis(side = 2, col.axis = "black", col = "black")
  legend("bottomright", legend = paste0("pretr = ", bestpretr), bty = "n", 
         text.col = "black")
  # legend("right", legend = levels(as.factor(datCal$DateDOY)), pch = 19, col = col.doy, bty = "n", text.col = "black")
  #valid
  datVal[[bestpretr]]$pred <-
    drop(predict(bestpretrmod$model, ncomp = bestpretrmod$model$ncomp, newdata = datVal[[bestpretr]]$NIRS))
  valid.out <- stat_pred(dataSet = datVal[[bestpretr]], obs = i, pred = "pred")
  plot(pred ~ get(i), datVal[[bestpretr]], pch = 21, bg = "gray", 
       xlab = "obs", ylab = "pred", main = paste(i, "(val)", sep = " "), 
       axe = F, col.main = "black", col.lab = "black")
  axis(side = 1, col.axis = "black", col = "black")
  axis(side = 2, col.axis = "black", col = "black")
  abline(0, 1, lty = 2, lwd = 2, col = "black")
  abline(valid.out$modlin, lty = 1, lwd = 2, col = "black")
  legend("topleft", bty = "n", c(
    # paste("R2_reg = ", round(valid.out$output[1], 2) , sep = ""),
    paste("R2_val = ", round(valid.out$output[2], 2) , sep = ""),
    paste("RMSE_val = ", round(valid.out$output[3], 2), sep = ""),
    # paste("RPD_val = ", round(valid.out$output[4], 2), sep = ""),
    paste("nobs = ", valid.out$output[5], sep = "")), 
    text.col = "black")
  #output
  out <- data.frame(i, bestpretr, length(bestpretrmod$model$model$trait), length(model$outliers[[bestpretr]]), t(bestpretrmod$output), t(valid.out$output))
  colnames(out) <- c("trait", "pretr", "n", "nb.out", "nb.comp", "R2.train", "R2.MCCV.mean", "R2.MCCV.sd", "RMSE.MCCV.mean", "RMSE.MCCV.sd", "RPD.MCCV.mean", "RPD.MCCV.sd", "R2.reg", "R2.val", "RMSE.val", "RPD.val", "n.val")
  kable(out[, c("trait", "pretr", "n", "nb.out", "nb.comp", "R2.train", "R2.MCCV.mean", "R2.MCCV.sd", "RMSE.MCCV.mean", "RMSE.MCCV.sd", "R2.reg", "R2.val", "RMSE.val", "n.val")])
  output.cal.KS <- rbind(output.cal.KS, out)
  mod.bestpretr[[i]] <- bestpretrmod
  #cleaning
  rm(datCal, bestpretr, datVal, out, valid.out, bestpretrmod, model)
}
```

### Best models

```{r}
kable(output.cal.KS[, c("trait", "pretr", "n", "nb.out", "nb.comp", "R2.MCCV.mean", "R2.MCCV.sd", "RMSE.MCCV.mean", "RMSE.MCCV.sd")], "html",
      row.names = FALSE, digits = 2) %>%
  kable_styling(bootstrap_options = "striped", full_width = F, font_size = 12)
```

### Save

```{r eval=FALSE}
setwd("D:/Mes Donnees/Doctorat/Axe 1 - NIRS/Model_final_propre_OK")
save(list = c("output.cal.KS", "mod.bestpretr"), file = "output_cal_pred_MicroNIR_KS.Rdata")
```

## Calibration with detection and elimination of outliers and with independent validation set (noout)

```{r cache = TRUE, fig.height=6, fig.width=12}
output.cal.noout.KS <- NULL
mod.bestpretr.noout <- NULL
outliers.KS <- NULL
for (i in trait.ids){
  #data prep
  datCal <- lapply(nirs.list.train, function(x){
    droplevels(na.omit(x[, c("echantillon", "DateDOY", "cépage", i, "NIRS")]))
  })
  #model
  model <- calibrationSPIR(calibSetList = datCal, chem = i, maxcomps = 20,
                           LOO = FALSE, folds = 5, iterMCCV = 50, wold = 1, thresh = 0.001)
  print(i)
  print(model$outcalib)
  #best pretr
  bestpretr <- rownames(model$outcalib)[which.max(model$outcalib$R2.MCCV.mean)]
  print(bestpretr)
  #best pretr mod
  datCal.noout <- subset(datCal[[bestpretr]], ! echantillon %in% model$outliers[[bestpretr]])
  bestpretrmod <- MCCV(dat = datCal.noout, trait = i,
                       maxcomp = 20, fold = 5,
                       iter = 50, crit = 1)
  bestpretrmod$output
  datVal <- lapply(nirs.list.valid.ok, function(x){
    droplevels(na.omit(x[, c("echantillon", "DateDOY", "cépage", i, "NIRS")]))
    })
  #plots
  par(mfrow = c(1, 2))
  #plot by DOY
  plotMCCV(bestpretrmod, pch = 19,
           col = col.doy[as.factor(datCal.noout$DateDOY)],
           main = paste(i, "(cv)", sep = " "), axe = F, col.main = "black", col.lab = "black")
  axis(side = 1, col.axis = "black", col = "black")
  axis(side = 2, col.axis = "black", col = "black")
  legend("bottomright", legend = paste0("pretr = ", bestpretr), bty = "n", text.col = "black")
  # legend("right", legend = levels(as.factor(datCal.noout$DateDOY)), pch = 19, col = col.doy, bty = "n", text.col = "black")
  #valid
  datVal[[bestpretr]]$pred <-
    drop(predict(bestpretrmod$model, ncomp = bestpretrmod$model$ncomp, newdata = datVal[[bestpretr]]$NIRS))
  valid.out <- stat_pred(dataSet = datVal[[bestpretr]], obs = i, pred = "pred")
  plot(pred ~ get(i), datVal[[bestpretr]], pch = 21, bg = "gray", xlab = "obs", ylab = "pred", main = paste(i, "(val)", sep = " "), col.main = "black", col.lab = "black", axe = F)
  axis(side = 1, col.axis = "black", col = "black")
  axis(side = 2, col.axis = "black", col = "black")
  abline(0, 1, lty = 2, lwd = 2, col = "black")
  abline(valid.out$modlin, lty = 1, lwd = 2, col = "black")
  legend("topleft", bty = "n", c(
    # paste("R2_reg = ", round(valid.out$output[1], 2) , sep = ""),
    paste("R2_val = ", round(valid.out$output[2], 2) , sep = ""),
    paste("RMSE_val = ", round(valid.out$output[3], 2), sep = ""),
    # paste("RPD_val = ", round(valid.out$output[4], 2), sep = ""),
    paste("nobs = ", valid.out$output[5], sep = "")
  ), text.col = "black")
  #output
  out <- data.frame(i, 
                    bestpretr, 
                    length(bestpretrmod$model$model$trait), 
                    length(model$outliers[[bestpretr]]), 
                    t(bestpretrmod$output), 
                    t(valid.out$output))
  
  colnames(out) <- c("trait", 
                     "pretr", 
                     "n", 
                     "nb.out", 
                     "nb.comp", 
                     "R2.train", 
                     "R2.MCCV.mean", 
                     "R2.MCCV.sd", 
                     "RMSE.MCCV.mean", 
                     "RMSE.MCCV.sd", 
                     "RPD.MCCV.mean", 
                     "RPD.MCCV.sd", 
                     "R2.reg", 
                     "R2.val", 
                     "RMSE.val", 
                     "RPD.val", 
                     "n.val")
  
  kable(out[, c("trait", 
                "pretr", 
                "n", 
                "nb.out", 
                "nb.comp", 
                "R2.train", 
                "R2.MCCV.mean", 
                "R2.MCCV.sd", 
                "RMSE.MCCV.mean", 
                "RMSE.MCCV.sd", 
                "R2.reg", 
                "R2.val", 
                "RMSE.val", 
                "n.val")])
  
  output.cal.noout.KS <- rbind(output.cal.noout.KS, out)
  
  mod.bestpretr.noout[[i]] <- list(bestpretrmod, bestpretr, out)
  
  #outliers
  outliers.KS[[i]] <- model$outliers
  
  #cleaning
  rm(datCal, bestpretr, datCal.noout, out, bestpretrmod, model)
}
```

### Best models

```{r}
kable(output.cal.noout.KS[, c("trait", "pretr", "n", "nb.out", "nb.comp", "R2.MCCV.mean", "R2.MCCV.sd", "RMSE.MCCV.mean", "RMSE.MCCV.sd")], "html",
      row.names = FALSE, digits = 2) %>%
  kable_styling(bootstrap_options = "striped", full_width = F, font_size = 12)
```

### Save

```{r eval=FALSE}
setwd("D:/Mes Donnees/Doctorat/Axe 1 - NIRS/Model_final_propre_OK")
save(list = c("output.cal.noout.KS", "outliers.KS", "mod.bestpretr.noout"), file = "output_cal_pred_MicroNIR_KS_noout.Rdata")
```
